---
title: "PS811 Portfolio"
author: "Monica Komer"
output:
  html_document: 
    toc: true ## table of contents
    code_folding: hide
---
```{r setup, include=FALSE}
## Defaults for R chunks
knitr::opts_chunk$set(echo = TRUE, ## echo = TRUE means code will show
                      warning=FALSE, ## supress warnings and messages from R
                      message=FALSE,
                      fig.path='Figs/', ## where to save figures
                      fig.height = 3, ## default figure size (can differ in each chunk)
                      fig.width = 3)

## Add any R packages you require. 
## Here are some we will use in 811:
requires <- c("tidyverse", ## tidyverse includes dplyr and ggplot2
              "magrittr",
              "foreign",
              "readstata13",
              "here")

## Install any you don't have
to_install <- c(requires %in% rownames(installed.packages()) == FALSE)
install.packages(c(requires[to_install], "NA"), repos = "https://cloud.r-project.org/" )

## Load all required R packages
library(tidyverse)
library(ggplot2); theme_set(theme_minimal())
library(magrittr)
library(here)
library(dplyr)
library(sjmisc)
library(sjlabelled)
library(sjPlot)
library(sjstats)
library(ggeffects)
library(texreg)
library(broom)
library(lme4)
```



### Research Interests
The overarching focus of my research is on the causes and consequences of women’s representation with a particular focus on the Middle East and North Africa. More specifically, I am interested in studying how (and which) women gain leadership positions in political parties and the effect of women’s presence on citizen’s attitudes and behaviors. 

### Questions I hope to answer
Which women gain leadership positions in secular and religious parties? What determines which women run, where (i.e. district location) they run, and their position on party lists?

Do female politicians interact with constituents differently than their male colleagues? Is this conditional on party affiliation or quota type? Are citizens more loyal to regimes that enact gender quotas? Are women more interested in politics when represented by a woman instead of a male?

---

### Methods
Various regression models including hierarchical models, survey experiments, text analysis, and qualitative methods (primarily interviews and focus groups). 

---

### Relevent data

- Arab Barometer ( [data](http://www.arabbarometer.org/survey-data/) )

- World Values Survey ( [data](http://www.worldvaluessurvey.org/WVSContents.jsp) )

- Afro Barometer ( [data](http://www.afrobarometer.org/data) )

- Quota Adoption and Reform over Time (QAROT) dataset ( [data](https://www.openicpsr.org/openicpsr/project/100918/version/V1/view) )

- Tunisia Election Data (results by district available [here](http://tunisiaelectiondata.com/) )


---

### Data Visualization

### Is there a gender gap in attitudes towards female politicians? Do attitudes vary by country?

Using Arab Barometer data to evaluate attitudes towards female leaders: 

```{r ABIV_data}
d <- read.csv(here("data/ABIV_English_subset.csv"))
``` 

I evaluate responses to the statement: "men are better at political leadership than women." I recode this variable:

```{r ABIV_vars}

#naming things and recoding leader variable (so higher values indicate more gender egalitarian views)
d$leader <- as.numeric(d$q6013)
d$leader[as.numeric(d$q6013) == 3] <- 2
d$leader[as.numeric(d$q6013) == 4] <- 3
d$leader[as.numeric(d$q6013) == 5] <- 1
d$leader[as.numeric(d$q6013) == 6] <- 4
d$leader[as.numeric(d$q6013) < 3] <- NA


d$female <- as.numeric(d$q1002)
d$female[as.numeric(d$q1002) == 2] <- 0



```
<!-- When I run the code you recommended: d$leader %<>% 
  recode(`3` = 2,
         `4` = 3,
         `5` = 1,
         `6` = 4,
         .default = NaN) 
or 
d$female %<>% recode(`1` = 1, `2`= 0)
I get an error: "No applicable method for 'recode' applied to an object of class "NULL".-->


Mean response by gender and country with higher values indicating more gender egalitarian views (i.e., less supportive of the statement that "men are better at political leadership than women"):
```{r means}

#Mean response by gender
d %>%
  group_by(female) %>% summarise(leader = mean(leader, na.rm=TRUE))



#Mean response by country
d %>%
  group_by(country) %>% summarise(leader = mean(leader, na.rm=TRUE))

```

Plotting responses by age and gender: 
```{r age-histogram, fig.width=8}
d %>% 
  mutate(age = as.numeric(q1001),
         gender = q1002) %>% 
  mutate(agree = ifelse(str_detect(q6013,"agree"), "Agree", NA)) %>%
  mutate(agree = ifelse(str_detect(q6013, "disagree"), "Disagree", agree)) %>%
  filter(!is.na(agree) ) %>%
  ggplot() +
  aes(x = age, fill = q6013) +
  labs(x= "Age", y = "Number of Respondents", fill = "Response", title= "Agreement with ''Men are better at political leadership'' ") +
  geom_histogram(binwidth = 10) + facet_grid(gender ~ agree)
```




```{r Country-Histogram, fig.width=8, eval=FALSE}
d %>%
  group_by(country) %>%
  mutate(leader_m = mean(leader, na.rm= TRUE)) %>%
  distinct(country, leader_m) %>%
  ggplot() + 
  aes(x = reorder(country, leader_m), y = leader_m) +
  labs(x= "Country", y="Mean Response", title= "Attitudes Towards Female Leaders") +
  geom_histogram(binwidth= 15, stat= "identity")
```


We can also visualize the mean response by gender and country (with larger values indicating more gender egalitarian attitudes):
```{r Gender-Histogram, fig.width=8}
d %>%
  mutate(gender = q1002) %>%
  group_by(country, gender) %>%
  mutate(leader_m = mean(leader, na.rm= TRUE)) %>%
  distinct(country, leader_m) %>%
  ggplot() + 
  aes(x = reorder(country, leader_m), y = leader_m, fill = gender) +
  labs(x= "Country", y="Mean Response", title="Attitudes Towards Female Leaders") +
  geom_histogram(binwidth=8, stat= "identity")
```


An alternative, and somewhat easier, way to visualize the gender gap in attitudes towards female leaders:
```{r Gender Gap-plot, fig.width=8}

d %>%
  mutate(gender = q1002) %>%
  group_by(country, gender) %>%
  mutate(leader_m = mean(leader, na.rm= TRUE)) %>%
  distinct(country, leader_m) %>%
  ggplot() +
  geom_point() +
  aes(y = reorder(country, leader_m), x = leader_m, color = gender) +
  labs(y=" " ,x="Mean Response\n(Least to Most Supportive)", title= " Gender Gap in Attitudes Towards Female Leaders", color = "Gender")
```



```{r Gap-plot, fig.width=8}
d %>%
  mutate(gender = q1002) %>%
  group_by(country, gender) %>%
  mutate(leader_m = mean(leader, na.rm= TRUE)) %>%
  distinct(country, leader_m) %>%
  ggplot() + 
  geom_point(aes(color=gender)) +
  aes(y = reorder(country, leader_m), x = leader_m) +
  labs(y= "Country", x="Mean Response") +
  geom_line()
```


---

### Project Presentation

### Effects of Women's Representation on Public Opinion

Using Arab Barometer survey data and data from Tunisia's 2014 parliamentary elections: 

```{r ABTun data}
abIV <- read.csv(here("data/ABIV_English_subset.csv"))
elect <- read.csv(here("Data/TUN2014PAR_Gov.csv"))
``` 

Merging and recoding the data:

```{r merging}
abIV <- read.csv(here("data/ABIV_English_subset.csv"))
elect <- read.csv(here("Data/TUN2014PAR_Gov.csv"))

abTun <- filter(abIV, country =="Tunisia")
abTun$governorate<- as.character(abTun$q1)

abtun<- full_join(abTun, elect, by= "governorate") 

#creating variable to measure gender egalitarianism
abtun$leader <- as.numeric(abtun$q6013)

abtun$leader %<>%  # reorder so higher values mean more gender egal
  recode('5' = 1, #strongly agree
         '3' = 2,
         '2' = 3,
         '4' = 4,
         '6' = 5,
         .default = NaN)


#Binary 
abtun$leaderdi <- abtun$leader
abtun$leaderdi %<>%
  recode('5' = 1,
         '4' = 1,
         '2' = 0,
         '1' = 0,
         .default = NaN)

#Gender Var
abtun$female <- as.numeric(abtun$q1002)

abtun$female %<>% 
  recode('1' = 1, 
         '2' = 0,
         .default = NaN)

``` 

Plotting average gender egalitarian attitudes (binary measure) by women's representation across disricts: 

```{r mean plot, fig.width=8, fig.height=4}
##Plotting women's rep and avg gender egalitarian attitude by governorate
abtun %>%
  group_by(governorate) %>%
  mutate(meanleaderdi = mean(leaderdi, na.rm= TRUE)) %>%
  distinct(per_wom, meanleaderdi, governorate) %>%
  ggplot() +
  aes(x = meanleaderdi, y = per_wom) +
  geom_point() +
  geom_text(aes(label= governorate, vjust = 1)) +
  theme_bw() +
  geom_smooth(method=lm, se=FALSE) +
  labs(title= "Figure 1: Gender Egalitarian Attitudes and Women's Representation by Governorate",
       x= "Average Gender Egalitarian Attitudes",
       y= "Percent of Female Representatives")
```

Logistic Regression with district-level fixed effects:

$$ logit(y)=B_0 + B_1 * Percent Women_j + B_2 * Head Women_j + B_3 * Female_{ij} + u_j $$

In this equation, the outcome variable (y) is the presence or absence of gender egalitarian attitudes. $B_0$ is the log-odds that $y=1$ when the remaining variables equal 0.  $B_1$ is the effect on log-odds of a one-unit increase in the percent of female representatives (i.e., a one percent increase) conditional on the other variables in the model and for individuals in the same district. $u_j$ is the effect of being in district $j$ on the log-odds that the outcome, gender egalitarian attitudes, equals 1.

```{r logit}
test1 <- (glmer(leaderdi ~ per_wom + headwom_di + female + (1|governorate), weights = wt,
                data= abtun, family=binomial(link="logit")))

summary(test1)


``` 

Plotting the Effect:
```{r odds ratio, fig.width=7}
plot_model(test1, show.values= TRUE, show.p = FALSE, vline.color = "red") +
  labs(title = "Estimated Effect on Gender Egalitarian Attitudes")
```

Predicted Probability of women's representation on gender egalitarian attitudes with confidence bounds:
```{r pred probability, fig.width = 8}
plot(ggpredict(test1, terms = c("per_wom"))) +
labs(title= "Predicted Probabilities of Gender Egalitarianism", 
     x= "Percent of Female Representatives",
     y= "Probability of Gender Egalitarian Attitudes")
```


Predicted Probability Conditional on District. Plotting this effect for a random set of 9 districts:
```{r pred prob, fig.width=8}
#predicted prob for random set of 7 disricts
plot_model(test1,type="pred",
           terms=c("per_wom","governorate [sample = 7]"),pred.type="re") +
  labs(title= "Predicted Probabilities of Gender Egalitarian Attitudes",
       x = "Percent of Female Representatives",
       y= "Probability of Gender Egalitarian Attitudes")
```
